Higgs Assisted Razor Search for Higgsinos at a 100 TeV $pp$ Collider

### 2019-12-17 main.tex
Addressed issue 4
Shufang's edits

The existence of dark matter provides unambiguous evidence of new physics
beyond the Standard Model (SM) of particle physics. Indeed, given that there is
far more dark matter in the Universe than there is baryonic matter, determining
its precise nature is one of the most exciting challenges in particle physics
today.  One promising group of candidates for dark matter is the class of stable
particles known as weakly interacting massive particles (WIMPs). These
particles arise naturally in many extensions of the SM that are motivated by
the need to stabilize the electroweak scale. One of the most heavily studied extensions
is the Minimal Supersymmetric Standard Model (MSSM), which predicts a
`supersymmetric partner', or \emph{superpartner}, for each SM
particle\footnote{For a detailed review of the MSSM, see Ref.
\cite{Martin:1997ns}.}. In the MSSM with \emph{R}-parity conservation, the
lightest supersymmetric partner (LSP) is predicted to be absolutely stable,
making it a good candidate for dark matter. The identity of the LSP is
determined by the mass hierarchy of the superpartners, which in turn depends on
the mechanism of supersymmetry breaking.  However, taking into account experimental
constraints and phenomenological considerations, the lightest of the MSSM
particles known as the neutralinos emerges as the most attractive candidate
for the LSP \cite{Bertone:2004pz}.

 % Why go beyond the MSSM?
Though a natural MSSM spectrum has the potential to tame the hierarchy problem,
it is under siege from recent data from the Large Hadron Collider (LHC)
\cite{Aaboud:2018ujj, Sirunyan:2018vjp}.  A compelling alternative scenario
comes in the form of split supersymmetry (split SUSY) \cite{Wells:2003tf,
ArkaniHamed:2004yi, Giudice:2004tc}. In this scenario, the lightest
superpartners are the fermionic ones, on the scale of 1-10 TeV, while the
scalar superpartners can be much heavier, on the scale of 100 - 1000 TeV. In
exchange for accepting some level of fine-tuning, we obtain numerous benefits,
including the suppression of flavor-changing neutral currents and greater
compatibility with data from CP-violation experiments.  The low lying fermionic
superpartners consist of neutralinos and charginos (collectively known as
electroweakinos), and gluinos. The current LHC search limits on gluinos are
already around 2 TeV for prompt decay and 2.5 TeV for disappearing track
searches~\cite{CMS-PAS-SUS-19-005}. The limits on electroweakinos with heavy
scalar superparticles, however, are still relatively weak: between \ap{0.15 $-$ 1.0
TeV, depending on the mixing among the gauge
eigenstates~\cite{ATL-PHYS-PUB-2019-022, Sirunyan:2017lae, Aaboud:2018htj,
Sirunyan:2018ubx, Aaboud:2018zeb}} \Adarsh{TODO: Confirm that these are the
latest limits.}, depending on the decay modes of the parent electroweakinos and
their species.  In this paper, we will focus on the discovery potential of the
neutralinos and charginos at a 100 TeV $pp$ collider.

@@

% Why a 100 TeV collider?
A 100 TeV $pp$ collider would open up an immense number of physics
opportunities not afforded to the 14 TeV LHC, including \st{exploring the most
interesting regions of the split SUSY parameter space} \ap{discovering thermal
electroweakino dark matter candidates.}
\cite{Arkani-Hamed:2015vfh}.
% Status of 100 TeV planning
Such a collider would be a natural next step after the LHC, and is being
actively discussed in the particle physics community, with two major proposals
being the Future Circular Collider (FCC)-hh by CERN \cite{FCC-hh}, and the
Super Proton Proton Collider (SppC)  in China \cite{CEPC}.  As the next energy
frontier machine, it is crucial to fully explore the physics
potential of a 100 TeV $pp$ collider, especially for new particles that are
either too heavy or too rare to be produced at the
LHC~\cite{Arkani-Hamed:2015vfh,Contino:2016spe,Golling:2016gvc,Mangano:2016jyj}.
In particular, studies have already been performed on the prospects of
discovering neutralino dark matter for compressed \cite{Low:2014cba,
diCortona:2014yua,Cirelli:2014dsa,Mahbubani:2017gjh,Mahbubani:2018tin,Han:2018wus} and
well-separated \cite{Gori:2014oua,Acharya:2014pua} neutralino spectra.

@@

The optimal search strategy for finding electroweakinos   highly depends on
their mass spectra \cite{Han:2013kza}.  Phenomenological collider studies on
finding electroweakinos with a small mass splitting of about 0.1 $-$ 50 GeV at a
100 TeV $pp$  collider can be found in \cite{Low:2014cba, Bramante:2014tba,
Berlin:2015aba, Cirelli:2014dsa}.
For $3\ \text{ab}^{-1}$ of data, the projected exclusion reaches for pure Wino
LSPs using the monojet channel is about $1.4$~TeV.  The reach
can be increased to $\sim3$ TeV with a disappearing tracks search.
Neutralinos up to 1 TeV can be excluded for mass splittings of about 20-30 GeV,
using a soft lepton search \cite{Low:2014cba}. Additionally, Bino-Wino LSPs
with masses up to 1.5 TeV for an inter-state mass splitting of around 1 GeV  can be
discovered at a 100 TeV $pp$ collider with $\sim7\ \text{ab}^{-1}$
of data \cite{Bramante:2014tba}.

Well-separated spectra have been studied in \cite{Gori:2014oua,
Acharya:2014pua} using multi-lepton channels, of the form

@@

For each signal mass combination and each background process, we preselected
events that passed the lepton trigger, contained two SFOS leptons and two
$b$-tagged jets, and calculated a number of kinematic variables for each event.   We then
placed these kinematic variables in an array, with each row corresponding to an
event, and the columns corresponding to the kinematic variables. A mixture of
low-level and high-level kinematic variables was shown to have the greatest
effectiveness. The kinematic variables chosen were $m_{\ell\ell}$, $m_{bb}$,
$M_R$, $M_T^R$, $\slashed{E}_T,$ the total hadronic transverse energy $H_T$,
and the transverse momenta of individual final state leptons and $b$-quarks:
$p_T^{\ell_1}$, $p_T^{\ell_2}$, $p_T^{b_1}$, and $p_T^{b_2}$. We then divided
the generated events into training and test sets, with the training set
comprising 75\% of signal events and 30\% of background events.  We used the
\texttt{scikit-learn} package  \cite{Pedregosa2011} to implement our analysis.
A boosted decision tree classifier was trained with 1000 weak learners and a
learning rate of 0.025.

@@

For certain values of the
cuts on kinematic variables or the score of the classifier, no events survived
for one or more of the background components. To guard against overly optimistic
estimates of the significance $S/\sqrt{B}$, we set a lower bound of 3 for the
number of Monte Carlo (MC) events after cuts for each background component, corresponding to
the 95\% confidence interval for a Poisson distribution.

>>>>

The existence of dark matter provides unambiguous evidence of new physics
beyond the Standard Model (SM) of particle physics. Indeed, given that there is
far more dark matter in the Universe than there is baryonic matter, determining
its precise nature is one of the most exciting challenges in particle physics
today.  One promising group of candidates for dark matter is the class of stable
particles known as weakly interacting massive particles (WIMPs). These
particles arise naturally in many extensions of the SM that are motivated by
the need to stabilize the electroweak scale. One of the most heavily studied extensions
is the Minimal Supersymmetric Standard Model (MSSM), which predicts a
`supersymmetric partner', or \emph{superpartner}, for each SM
particle\footnote{For a detailed review of the MSSM, see Ref.
\cite{Martin:1997ns}.}. In the MSSM with \emph{R}-parity conservation, the
lightest supersymmetric partner (LSP) is predicted to be absolutely stable,
making it a good candidate for dark matter. The identity of the LSP is
determined by the mass hierarchy of the superpartners, which in turn depends on
the mechanism of supersymmetry breaking.  However, taking into account experimental
constraints and phenomenological considerations, \st{the lightest of the MSSM
particles known as the neutralinos emerges as the most attractive candidate
for the LSP}
\Shufang{neutralino LSP emerges as the most attractive candidate for WIMP dark matter.}~\cite{Bertone:2004pz}.

 % Why go beyond the MSSM?
Though a natural MSSM spectrum has the potential to tame the hierarchy problem,
it is under siege from recent data from the Large Hadron Collider (LHC)
\cite{Aaboud:2018ujj, Sirunyan:2018vjp}.  A compelling alternative scenario
comes in the form of split supersymmetry (split SUSY) \cite{Wells:2003tf,
ArkaniHamed:2004yi, Giudice:2004tc}. In this scenario, the lightest
superpartners are the fermionic ones, on the scale of 1-10 TeV, while the
scalar superpartners can be much heavier, on the scale of 100 - 1000 TeV. In
exchange for accepting some level of fine-tuning, we obtain numerous benefits,
including the suppression of flavor-changing neutral currents and greater
compatibility with data from CP-violation experiments.  The low lying fermionic
superpartners consist of neutralinos and charginos (collectively known as
electroweakinos), and gluinos. The current LHC search limits on gluinos are
already around 2 TeV for prompt decay and 2.5 TeV for disappearing track
searches~\cite{CMS-PAS-SUS-19-005}. \Shufang{The limits on electroweakinos with heavy scalar superparticles, however, are still relatively weak: between 150 GeV $-$ 1.0 TeV, depending on the neutralino/chargino species, decay modes, and their mass splittings~\cite{ATL-PHYS-PUB-2019-022, Sirunyan:2017lae,Sirunyan:2018ubx,  Aaboud:2018htj,Aaboud:2018zeb, 
Aad:2019qnd,Aad:2019vvf,Aad:2019vnb,Aaboud:2018ngk,Aaboud:2018sua,Aaboud:2018jiw,Aaboud:2017leg,Aaboud:2017mpt,Aaboud:2017nhr}.  In this paper, we will focus on the discovery potential of the
neutralinos and charginos at a 100 TeV $pp$ collider. }
{\bf Can not find Ref. 9}

@@

% Why a 100 TeV collider?
A 100 TeV $pp$ collider would open up an immense number of physics
opportunities not afforded to the 14 TeV LHC. \st{including exploring the most
interesting regions of the split SUSY parameter space}~\cite{Arkani-Hamed:2015vfh}.
% Status of 100 TeV planning
Such a collider would be a natural next step after the LHC, and is being
actively discussed in the particle physics community, with two major proposals
being the Future Circular Collider (FCC)-hh by CERN \cite{FCC-hh}, and the
Super Proton Proton Collider (SppC)  in China \cite{CEPC}.  As the next energy
frontier machine, it is crucial to fully explore the physics
potential of a 100 TeV $pp$ collider, especially for new particles that are
either too heavy or too rare to be produced at the
LHC~\cite{Arkani-Hamed:2015vfh,Contino:2016spe,Golling:2016gvc,Mangano:2016jyj}.
In particular, studies have already been performed on the prospects of
discovering neutralino dark matter for compressed \cite{Low:2014cba,Cirelli:2014dsa,Mahbubani:2017gjh,Mahbubani:2018tin,Han:2018wus,Bramante:2014tba, Berlin:2015aba, 
Fukuda:2017jmk, Saito:2019rtg} and
well-separated \cite{Gori:2014oua,Acharya:2014pua,diCortona:2014yua} neutralino spectra.

@@

\Shufang{
The optimal search strategy for finding electroweakinos is highly dependent on
their mass spectra \cite{Han:2013kza}. There are phenomenological collider studies on
finding electroweakinos with a small mass splitting of about 0.1 $-$ 50 GeV at the LHC~\cite{Giudice:2010wb, Gori:2013ala, Han:2013usa, Schwaller:2013baa, Baer:2014cua, Han:2014kaa} and 
at a 100 TeV $pp$  collider~\cite{Low:2014cba, Bramante:2014tba,
  Berlin:2015aba, Cirelli:2014dsa, Mahbubani:2017gjh, Mahbubani:2018tin,
Fukuda:2017jmk, Saito:2019rtg,Han:2018wus}.
For 100 TeV $pp$ collider with $30\ \text{ab}^{-1}$ of data, disappearing track search can discover Wino (Higgsino) LSP with mass up to about 4.7 TeV (1 TeV)~\cite{Low:2014cba,Cirelli:2014dsa,Mahbubani:2017gjh,Saito:2019rtg,Han:2018wus}.   The discovery reach from Monojet, mono-$Z$, mono-photon search is weaker: about 1 TeV (600 GeV) for Wino (Higgsino)~\cite{Low:2014cba,Cirelli:2014dsa,Mahbubani:2018tin,Han:2018wus}.  For mass splittings of about 20-30 GeV between low-lying neutralinos and charginos, soft lepton plus missing $E_T$ signal can be used~\cite{Low:2014cba,Bramante:2014tba}.  In addition, Weak boson fusion process can discover a Wino (Higgsino) up to about 1 TeV (530 GeV)~\cite{Cirelli:2014dsa, Berlin:2015aba}.  Note, however,  that in certain
scenarios, light Higgsinos between 75 $-$ 100 GeV are still viable, evading LEP and LHC
constraints \cite{Egana-Ugrinovic:2018roi}.}

Well-separated spectra have been studied using multi-lepton channels
\cite{Gori:2014oua, Acharya:2014pua}, of the form

@@

For each signal mass combination and each background process, we preselected
events that passed the lepton trigger, contained two SFOS leptons and two
$b$-tagged jets, and calculated a number of kinematic variables for each event.   We then
placed these kinematic variables in an array, with each row corresponding to an
event, and the columns corresponding to the kinematic variables. A mixture of
low-level and high-level kinematic variables was shown to have the greatest
effectiveness. The kinematic variables chosen were $m_{\ell\ell}$, $m_{bb}$,
$M_R$, $M_T^R$, $\slashed{E}_T,$ the total hadronic transverse energy $H_T$,
and the transverse momenta of individual final state leptons and $b$-quarks:
$p_T^{\ell_1}$, $p_T^{\ell_2}$, $p_T^{b_1}$, and $p_T^{b_2}$. We then divided
the generated events into training and test sets, with the training set
comprising 75\% of signal events and 30\% of background events.  \Shufang{More details here -- Referee comment 9.} We used the
\texttt{scikit-learn} package  \cite{Pedregosa2011} to implement our analysis.
A boosted decision tree classifier was trained with 1000 weak learners and a
learning rate of 0.025.

@@

For certain values of the
cuts on kinematic variables or the score of the classifier, no events survived
for one or more of the background components. To guard against overly optimistic
estimates of the significance $S/\sqrt{B}$, we set a lower bound of 3 for the
number of Monte Carlo (MC) events after cuts for each background component, corresponding to
the 95\% confidence interval for a Poisson distribution.  \Shufang{SS: Is it 3 MC events, or 3 weighted events after CS values used? Referee comment 10.}
